---
title: "Final assignment"
output: html_document 
---
##Data, year and variables and research question: does x, y, c and b affect one's satisfaction in life?  
## bases the knowledfe of Martha Nussbaums list of capabilities that affects people's well-being: life satisfaction as indicator of subjective well-being and the other variables indicators of objective (capability) well-being factors. 
JHLLT Does your health in any way limit your daily activities compared to most people of your age?
JLKMOVE If you could choose, would you stay here in your present home or would you prefer to move somewhere else?
JHSPRBQ Does your accommodation have any of the following problems? Vandalism or crime in the area
JQFEDHI Education
JGHQD Have you recently felt capable of making decisions about things? 
JGHQF Have you recently felt you couldn’t overcome your difficulties?
JGHQH Have you recently been able to face up to problems?
JHSCANB Here is a list of things which people might have or do. Please look at this card and tell me which things you (and your household) have or do? Pay for a week’s annual holiday away from home.
JHSCNTB Would you like to be able to pay for a week’s annual holiday away from home, but must do without because you cannot afford it?
JHSCAND Here is a list of things which people might have or do. Please look at this card and tell me which things you (and your household) have or do? Buy new, rather than second hand, clothes.
JHSCNTD Would you like to be able to buy new, rather than second hand, clothes, but must do without because you cannot afford it?
JHSCANF Here is a list of things which people might have or do. Please look at this card and tell me which things you (and your household) have or do? Have friends or family for a drink or meal at least once a month
JGHQC Have you recently felt that you were playing a useful part in things?
JVOTE7 Did you vote in this (past) year’s general election?
JLFSATO How dissatisfied or satisfied are you with your life overall?


```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(haven, ggplot2,knitr, pander,descr,foreign,dplyr,Hmisc,psych,GPArotation,plyr, dplyr, rmarkdown)
BHPS <- read.csv(file = "/Applications/IODS-project/data/BHPS.csv", sep=",", header =TRUE)
BHPS$X <- NULL
```
## Information about the data
```{r}
dim(BHPS)
describe(BHPS)
```


#### Information about the variables
```{r}
summary(BHPS)
```


Next I will present a graphical overview of the dataset "learning2014". The graph below gives hints of correlations between variables. According to the graph, it seems that there are rather little correlations between the learning approaches, gender and age with the exam points. Contrary, the strongest correlation is with the atttitude and points. 

```{r, echo= FALSE}

library(GGally)
library(ggplot2)

pairs(BHPS)

p <- ggpairs(BHPS)
p

ggpairs(BHPS, lower = list(combo = wrap("facethist", bins = 20)))
```


####Based on the graphical overview of the data, it seems that attitude correlates greatly with points. Let's take a closer look of all the variables:

#### ### Regression analysis: relationship of attitude, deep learning approach, age, and exam points as a dependent
####Regression model and analysis is probably one of hte most used method for predicting correlations between variables - to understand how certain independent variables (in my model that is attitude, deep learning approach and age) are related to the dependent variable (that is in this model examp points). Regression analysis helps to investigate different dimensions of the relationship and correlations. However, it is good to keep in mind that correlation does not (always) mean causality. (Armstrong, 2012.) The model1 (below) shows the correlations of the  independent variable and dependent variable.
#First linear regression model with all the variables - and then the ones that had the significant results 
```{r, echo= FALSE}
Model1 <- lm(JLFSATO~JVOTE7+JGHQC+JHSCANF+JHSCNTD+JHSCAND+JHSCNTB+JHSCANB+JGHQH+JGHQF+JGHQD+JHSPRBQ+JLKMOVE+JHLLT , data = BHPS)
summary(Model1)
```

### Interesting enough, only five variables seems to have significant correlation with life satisfaction. No other variables (i.e. any learning approaches, age nor gender) have statisctical signifance with the students’ exam points other than attitude. Thus, since there is no statistical significanse between age and deep learning approach, I will remove those two variables, and create a model2, with only attitude as independent variable and exam point as dependent variable. Although should we ignore the variables that does not have statistical significance, perhaps not. In the articel “Statistical errors. P values, ‘the gold standard’ of statistical validity, are not as reliable as many scientist assume”, Regina Nuzzo summarizes how nowadays P value of 0.01 means for a lot of scientists that there is just a 1% of chance that the result is being a false. However, p value does not tell that. All what P value is doing, is summarizing the data by asuuming a spexifix null hypothesis (=that there is no correaltion between these variables, but our world does not work like that, usually variables do correlate). (Nuzzo, 2014.)

##P value tells us yet something about existing correlation. Therefore, let’s take a closer look at regression model “model2”, where attitude towards statistics is an independent variable (since it was the only one with a statistical significance) and exam points a dependent variable:
## significant factors JHGQC = experience of valuable/useful, JHSCANF = dinner/drink with friends/familu, JGHQF = capability to overcome difficultie, JLKMOVE = capability to move, JHLLT = health limits

```{r}
Model2 <- lm(JLFSATO~JGHQC+JHSCANF+JGHQF+JLKMOVE+JHLLT , data = BHPS)
summary(Model2)
```

### then some visualization JHGQC = experience of valuable/useful, JHSCANF = dinner/drink with friends/familu, JGHQF = capability to overcome difficultie, JLKMOVE = capability to move, JHLLT = health limits

```{r, echo=FALSE}
library(ggplot2)
plot(Model2, 1, par(mfrow = c(2,2)))
plot(Model2, 2, par(mfrow = c(2,2)))
plot(Model2, 5, par(mfrow = c(2,2)))

plot(Model2, which = c(1,2,5), par(mfrow = c(2,2)))
```

```{r}

qplot(BHPS$JGHQC, BHPS$JLFSATO) + geom_smooth(method = "lm")

ggplot(BHPS, aes(JLFSATO) + geom_errorbar(ymin=0.03, ymax=0.06, width=.2) + ylim(0, 1) +
geom_segment(x = 0, xend = 1, y = 0.52 + 0.01, yend = 0.52 + 0.01*1, colour="green") +    xlim(0, 1) +  geom_errorbar(ymin=0.01, ymax=0.03, width=.2) + ylim(0, 1)

ggplot(BHPS, aes(JLFSATO)) + stat_summary(fun.y=mean, geom="line") + ylab("predicted life satisfcation") + ylim(0,1) +  
  geom_segment(x = 0, xend = 1, y = 0.52 + 0.01, yend = 0.52 + 0.01*1, colour="green") +    xlim(0, 1) 

p1 <- ggplot(BHPS, aes(x = BHPS$JLFSATO, y = BHPS$JGHQF,  col = BHPS$JHLLT))
p2 <- p1 + geom_point()
p3 <- p2 + geom_smooth(method = "lm")
p4 <- p3 + ggtitle("Student's attitude versus exam points")
print(p4)

p5 <- ggplot(learning2014, aes(x = surf, y = Age, col = gender))
p6 <- p5 + geom_point()
p7 <- p6 + geom_smooth(method = "lm")
p8 <- p7 + ggtitle("Student's learning approach (surface) versus age")
p8
```



## and conclusions

## References 



